---
title: "Appendix"
author: "Max Aantjes"
date: "26/08/2020"
output:
  github_document:
    toc: true
    toc_depth: 2
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(openxlsx)
library(stringr)
library(tidyverse)
```


```{r}
replace_special_char <- function(x) {
        original <- c("á", "é", "ó", "ú", "ñ", "í")
        new <- c("a", "e", "o", "u", "n", "i")
        for(i in 1:length(original)) {
                x <- gsub(original[i], new[i], x)}
        return(x)}

df <- read.xlsx("Registros-Administrativos-2019-2020-Inicio.xlsx", sheet = "Registros Administrativos", startRow = 13)

names(df) <- tolower(replace_special_char(names(df)))
names(df)[2] <- "duplicate"
names(df)[34] <- "tot.students"

df <- df %>%
        select(-duplicate) %>%
        mutate(canton = factor(canton))  %>%
        mutate(sostenimiento = factor(sostenimiento, levels = c("Fiscal", "Fiscomisional", "Particular", "Municipal"), labels = c("public", "mixed", "private", "public"))) %>%
        filter(tot.students != 0) %>%
        filter(total.docentes != 0)
```

```{r}
df2 <- df %>%
        mutate(class.size = tot.students/total.docentes) 

p <- ggplot(data = df2, aes(x = class.size, y = sostenimiento))
p1 <- p + geom_boxplot() + facet_grid(~zona.inec)
p1
```


```{r}
df1 <- df %>%
        group_by(canton, sostenimiento) %>%
        summarise(n = sum(tot.students)) %>%
        pivot_wider(names_from = sostenimiento, values_from = n) %>%
        mutate(public = ifelse(is.na(public), 0, public)) %>%
        mutate(mixed = ifelse(is.na(mixed), 0, mixed)) %>%
        mutate(private = ifelse(is.na(private), 0, private)) %>%
        mutate(total = public + mixed + private) %>%
        mutate(public.percent = public/total) %>%
        mutate(mixed.percent = mixed/total) %>%
        mutate(private_percent = private/total)

df2 <- df1 %>%
        arrange(-private_percent) %>%
        mutate(private_percent = private_percent * 100) %>%
        select(canton, private_percent, total)

head(df2, 10)
```

```{r}
df1 <- df %>%
        group_by(zona, sostenimiento) %>%
        summarise(n = sum(tot.students)) %>%
        pivot_wider(names_from = sostenimiento, values_from = n) %>%
        mutate(public = ifelse(is.na(public), 0, public)) %>%
        mutate(mixed = ifelse(is.na(mixed), 0, mixed)) %>%
        mutate(private = ifelse(is.na(private), 0, private)) %>%
        mutate(total = public + mixed + private) %>%
        mutate(public.percent = public/total) %>%
        mutate(mixed.percent = mixed/total) %>%
        mutate(private_percent = private/total)

df2 <- df1 %>%
        arrange(-private_percent) %>%
        mutate(private_percent = private_percent * 100) %>%
        select(canton, private_percent, total)

head(df2, 10)   
```

```{r}
df1 <- df %>%
        group_by(sostenimiento) %>%
        summarise(no.teachers = sum(total.docentes))
df1
```

```{r}
df1 <- df %>%
        group_by(sostenimiento) %>%
        count()
df1
```

```{r}
y2009s <- "https://educacion.gob.ec/wp-content/plugins/download-monitor/download.php?id=10555"
y2010s <- "https://educacion.gob.ec/wp-content/plugins/download-monitor/download.php?id=10557"
y2011s <- "https://educacion.gob.ec/wp-content/plugins/download-monitor/download.php?id=10559"
y2012s <- "https://educacion.gob.ec/wp-content/plugins/download-monitor/download.php?id=10561"
y2013s <- "https://educacion.gob.ec/wp-content/plugins/download-monitor/download.php?id=12176"
y2014s <- "https://educacion.gob.ec/wp-content/plugins/download-monitor/download.php?id=10567"
y2015s <- "https://educacion.gob.ec/wp-content/plugins/download-monitor/download.php?id=10569"
y2016s <- "https://educacion.gob.ec/wp-content/plugins/download-monitor/download.php?id=10573"
y2017s <- "https://educacion.gob.ec/wp-content/plugins/download-monitor/download.php?id=12645"
y2018s <- "https://educacion.gob.ec/wp-content/plugins/download-monitor/download.php?id=15711"
y2019s <- "https://educacion.gob.ec/wp-content/plugins/download-monitor/download.php?id=15713"

links1 <- list(y2009s, y2010s, y2011s, y2012s, y2013s, 
               y2014s, y2015s, y2016s, y2017s)
links2 <- list(y2018s, y2019s)
 
load.xlsx <- function(x, y, z){
        y <- read.xlsx(x, sheet = y, startRow = z)
        return(y)}

dat1 <- load.xlsx(y2009s, y= 1, z = 11)

dfB <- read.xlsx(y2018s, sheet = 2, startRow = 13)
dfC <- read.xlsx(y2019s,  sheet = "Registros Administrativos", startRow = 13)
listA <- lapply(links1, load.xlsx, y = 1, z = 11)

dfA <- do.call(rbind, listA)
names(dfA) <- replace_special_char(names(dfA))
dfA1 <- dfA %>%
        select(Periodo, Total.Estudiantes, Total.Docentes, 
               Nombre.Institucion, Zona.Inec, Sostenimiento, Canton, Provincia)

names(dfB) <- replace_special_char(names(dfB))
dfB1 <- dfB %>%
        select(Periodo, Total.Estudiantes, Total.Docentes, 
               Nombre.Institucion, Zona.Inec, Sostenimiento, Canton, Provincia)

names(dfC) <- replace_special_char(names(dfC))
names(dfC)[2] <- "duplicate"
dfC1 <- dfC %>%
        select(-duplicate) %>%
        select(Periodo, "Total.Estudiantes" = 'Total.estudiantes.(Inicial.a.3ro..De.Bachillerato)', 
               Total.Docentes, Nombre.Institucion, Zona.Inec, Sostenimiento, Canton, Provincia)

df <- rbind(dfA1, dfB1, dfC1)
saveRDS(df, "data.rds")
```

```{r}
df <- readRDS("data.rds")
df1 <- df %>%
        select("year" = Periodo, "tot.students" = Total.Estudiantes, 
               "tot.teachers" = Total.Docentes, Sostenimiento, "name" = Nombre.Institucion,
               "area" = Zona.Inec, Canton, Provincia) %>%
        mutate(year = str_extract(year, "^[0-9]{4}-[0-9]{4}"))  %>%
        #mutate(year = as.Date(paste0(year, "-01-01"))) %>%
        mutate(finance.source= factor(Sostenimiento, levels = c("Fiscal", "Fiscomisional", 
                                                                "Particular", "Municipal",
                                                                "Particular Laico",
                                                                "Particular Religioso"), 
                                      labels = c("public", "mixed", "private", "public",
                                                 rep("private", 2)))) %>%
        mutate(area = str_remove(area, "INEC")) %>%
        mutate(canton = replace_special_char(Canton)) %>%
        filter(tot.students != 0 & tot.teachers != 0)

df2 <- df1 #%>%
        #filter(!canton %in% c("SAMBORONDON", "GUAYAQUIL", "DURAN", "RUMINAHUI", "DAULE", "QUITO", "SALINAS"))

dfCS <- df2 %>%
        filter(finance.source == "public") %>%
        mutate(class.size = tot.students/tot.teachers) %>%
        group_by(year, canton, area) %>%
        summarise(class.size = median(class.size))

missing_to_zero <- function(var){return(ifelse(is.na(var), 0, var))}

df3<- df2 %>%
        group_by(year, finance.source, canton, area) %>%
        summarise(n = sum(tot.students)) %>%
        pivot_wider(names_from = finance.source, values_from = n) %>%
        mutate(public = missing_to_zero(public)) %>%
        mutate(mixed = missing_to_zero(mixed)) %>%
        mutate(private = missing_to_zero(private)) %>%
        mutate(prop = (private/sum(public, mixed, private))*100) %>%
        left_join(dfCS, by = c("year", "canton", "area")) %>%
        ungroup() %>%
        mutate(area = factor(area))

df4 <- df3 %>%
        filter(year %in% c("2010-2011", "2013-2014", "2016-2017", "2019-2020"))# %>%
        #filter(prop != 0)

p <- ggplot(dat = df4, aes(y = class.size, x = prop)) + theme_bw() 
p1 <- p + geom_point(alpha = 1/2) + facet_grid(area~year)
p2 <- p1 + geom_smooth(method = "lm", formula = y~x) 
p3 <- p2 + theme(axis.text.x = element_text(angle = 90)) + labs(
  x = "Proportion of students enrolled\nin private education (%)", 
  y = "Median of the average-class-size in public schools",
  title = "Relationship between private education attendance\nand public school class sizes in Ecuadorian cantons",
  subtitle = "Data reflects student enrolment and teacher positions at the start of the school year in all\nprimary and secondary education institutions. The average-class-size was first calculated\nfor each school. The median average-class-size was then taken at canton level.", 
  caption = "AMIE Data collected by INEC (2010-2019)")
p3

## This suggests that private institutions are concentrated in areas where class sizes are already large. 
df5 <- df4 %>%
  filter(area == "Urbana" & year == "2019-2020")
r <- lm(formula = prop ~ class.size, data = df5)
summary(r)
plot(r)

### The Bresusch-Pagan Test
## It involves using a variance function and using a χ2-test to test the null hypothesis that heteroskedasticity is not present (i.e. homoskedastic) against the alternative hypothesis that heteroskedasticity is present.

df5$resi <- r$residuals
var.func <- lm(resi^2 ~ class.size, data = df5)
summary(var.func)
qchisq(.95, df = 1)
```

```{r}
name_check <- function(x, y){
  for(i in 1:length(names(x))){
    if(isFALSE(identical(names(x)[i], names(y)[i]))){
      sapply(c(i, names(x)[i], names(y)[i]), print)}}
  message("all other colnames identical")}
```

```{r}
## Sources
"https://blogs.iadb.org/educacion/es/covid-19-y-la-transicion-de-la-educacion-privada-a-la-publica-en-ecuador/"
"https://www.elnuevosiglo.com.co/articulos/07-2020-la-transicion-de-la-educacion-privada-la-publica"
"https://www.infobae.com/america/mexico/2020/08/04/la-sep-esta-lista-para-migracion-de-alumnos-de-escuelas-privadas-a-publicas/"
"https://www.infobae.com/america/mexico/2020/08/04/la-sep-esta-lista-para-migracion-de-alumnos-de-escuelas-privadas-a-publicas/"
"https://www.24-horas.mx/2020/08/24/educacion-en-tiempos-de-covid-crece-migracion-de-escuelas-privadas-a-publicas/"
"https://www.eldiariodechihuahua.mx/local/temen-migracion-de-escuelas-privadas-a-publicas-20200805-1692951.html"
"https://www.milenio.com/politica/sep-lista-migracion-alumnos-escuelas-privadas-publicas"
"https://politica.expansion.mx/mexico/2020/08/10/el-covid-19-pone-en-riesgo-a-1-de-cada-4-escuelas-privadas"
"https://www.elfinanciero.com.mx/empresas/advierten-que-4-de-cada-10-de-escuelas-privadas-cerraran-por-educacion-a-distancia"
```


Context

Prop of priv/public schools, prop of priv/public teachers, boxplot class sizes. 
story: private schools have smaller class sizes than public schools. 

Inter American Bank data + Mexico
Story: migration from private schools to public schools. If situation does not change class sizes in public institutions will grow.

Is there historical evidence?
We expect it to be a static model instead of a lag model. 
Test for Autoregressive models of different orders. 

## Clean data set
```{r}
df <- readRDS("data.rds")
df1 <- df %>%
        select("year" = Periodo, "tot.students" = Total.Estudiantes, 
               "tot.teachers" = Total.Docentes, Sostenimiento, "name" = Nombre.Institucion,
               "area" = Zona.Inec, "canton" = Canton, "province" = Provincia) %>%
        mutate(year = str_extract(year, "^[0-9]{4}"))  %>%
        mutate(year = as.Date(paste0(year, "-05-01"))) %>%
        mutate(finance.source = factor(Sostenimiento, levels = c("Fiscal", "Fiscomisional", 
                                                                "Particular", "Municipal",
                                                                "Particular Laico",
                                                                "Particular Religioso"), 
                                      labels = c("public", "mixed", "private", "public",
                                                 rep("private", 2)))) %>%
        select(-Sostenimiento) %>%
        mutate(area = factor(str_remove(area, "INEC"), 
               levels = c("Rural", "Urbana"), labels = c("rural", "urban"))) %>%
        mutate(province = factor(province)) %>%
        #group_by(province) %>% ## Because some cantons in different provinces have the same names
        #mutate(canton = factor(replace_special_char(as.character(canton)))) %>%
        filter(tot.students != 0 & tot.teachers != 0) #%>%
        #ungroup()
```

# Country level time-series regression
## Calculate Parameters at Country-level 

```{r}
df2 <- df1 %>%
        filter(finance.source == "public") %>%
        mutate(public.school.class.size = tot.students/tot.teachers) %>%
        group_by(year) %>%
        summarise(median.public.school.class.size = median(public.school.class.size))

missing_to_zero <- function(var){return(ifelse(is.na(var), 0, var))}

df3<- df1 %>%
        group_by(year, finance.source) %>%
        summarise(n = sum(tot.students)) %>%
        pivot_wider(names_from = finance.source, values_from = n) %>%
        mutate(public = missing_to_zero(public)) %>%
        mutate(mixed = missing_to_zero(mixed)) %>%
        mutate(private = missing_to_zero(private)) %>%
        mutate(proportion.private.students = (private/sum(public, mixed, private))*100) %>%
        left_join(df2, by = "year") %>%
        ungroup() 
```

step 1
Convert data to an object of class **ts**
```{r}
## 
## No package required
tsdf <- ts(df4, 
           start = c(2007, 1), # End date
           end = c(2019, 1), # Start date
           frequency = 1) # Number of observations per unit time
```

```{r}

df

df3 <- df2 %>%
        filter(finance.source == "public") %>%
        mutate(class.size = tot.students/tot.teachers) %>%
        group_by(year, canton, area) %>%
        summarise(class.size = median(class.size))

missing_to_zero <- function(var){return(ifelse(is.na(var), 0, var))}

df4<- df2 %>%
        group_by(year, finance.source, canton, area) %>%
        summarise(n = sum(tot.students)) %>%
        pivot_wider(names_from = finance.source, values_from = n) %>%
        mutate(public = missing_to_zero(public)) %>%
        mutate(mixed = missing_to_zero(mixed)) %>%
        mutate(private = missing_to_zero(private)) %>%
        mutate(prop = (private/sum(public, mixed, private))*100) %>%
        left_join(dfCS, by = c("year", "canton", "area")) %>%
        ungroup() %>%
        mutate(area = factor(area))


p <- ggplot(dat = df4, aes(y = class.size, x = prop)) + theme_bw() 
p1 <- p + geom_point(alpha = 1/2) + facet_grid(area~year)
p2 <- p1 + geom_smooth(method = "lm", formula = y~x) 
p3 <- p2 + theme(axis.text.x = element_text(angle = 90)) + labs(
  x = "Proportion of students enrolled\nin private education (%)", 
  y = "Median of the average-class-size in public schools",
  title = "Relationship between private education attendance\nand public school class sizes in Ecuadorian cantons",
  subtitle = "Data reflects student enrolment and teacher positions at the start of the school year in all\nprimary and secondary education institutions. The average-class-size was first calculated\nfor each school. The median average-class-size was then taken at canton level.", 
  caption = "AMIE Data collected by INEC (2010-2019)")
p3

## This suggests that private institutions are concentrated in areas where class sizes are already large. 
df5 <- df4 %>%
  filter(area == "Urbana" & year == "2019-2020")
r <- lm(formula = prop ~ class.size, data = df5)
summary(r)
plot(r)

### The Bresusch-Pagan Test
## It involves using a variance function and using a χ2-test to test the null hypothesis that heteroskedasticity is not present (i.e. homoskedastic) against the alternative hypothesis that heteroskedasticity is present.

df5$resi <- r$residuals
var.func <- lm(resi^2 ~ class.size, data = df5)
summary(var.func)
qchisq(.95, df = 1)
```

```{r}
GDPGR_AR2 <- dynlm(ts(GDPGR_level) ~ L(ts(GDPGR_level)) + L(ts(GDPGR_level), 2))
```

Why is it bad?
Private schools concentrated in particular areas (inter American bank data). Furthermore, private schools tend to concentrate in areas where there public school classes are large. 




