---
title: "Appendix"
author: "Max Aantjes"
date: "26/08/2020"
output:
  github_document:
    toc: true
    toc_depth: 2
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(openxlsx)
library(stringr)
library(tidyverse)
library(xts)
```


```{r}
replace_special_char <- function(x) {
        original <- c("á", "é", "ó", "ú", "ñ", "í")
        new <- c("a", "e", "o", "u", "n", "i")
        for(i in 1:length(original)) {
                x <- gsub(original[i], new[i], x)}
        return(x)}

df <- read.xlsx("Registros-Administrativos-2019-2020-Inicio.xlsx", sheet = "Registros Administrativos", startRow = 13)

names(df) <- tolower(replace_special_char(names(df)))
names(df)[2] <- "duplicate"
names(df)[34] <- "tot.students"

df <- df %>%
        select(-duplicate) %>%
        mutate(canton = factor(canton))  %>%
        mutate(sostenimiento = factor(sostenimiento, levels = c("Fiscal", "Fiscomisional", "Particular", "Municipal"), labels = c("public", "mixed", "private", "public"))) %>%
        filter(tot.students != 0) %>%
        filter(total.docentes != 0)
```

```{r}
df2 <- df %>%
        mutate(class.size = tot.students/total.docentes) 

p <- ggplot(data = df2, aes(x = class.size, y = sostenimiento))
p1 <- p + geom_boxplot() + facet_grid(~zona.inec)
p1
```


```{r}
df1 <- df %>%
        group_by(canton, sostenimiento) %>%
        summarise(n = sum(tot.students)) %>%
        pivot_wider(names_from = sostenimiento, values_from = n) %>%
        mutate(public = ifelse(is.na(public), 0, public)) %>%
        mutate(mixed = ifelse(is.na(mixed), 0, mixed)) %>%
        mutate(private = ifelse(is.na(private), 0, private)) %>%
        mutate(total = public + mixed + private) %>%
        mutate(public.percent = public/total) %>%
        mutate(mixed.percent = mixed/total) %>%
        mutate(private_percent = private/total)

df2 <- df1 %>%
        arrange(-private_percent) %>%
        mutate(private_percent = private_percent * 100) %>%
        select(canton, private_percent, total)

head(df2, 10)
```

```{r}
df1 <- df %>%
        group_by(zona, sostenimiento) %>%
        summarise(n = sum(tot.students)) %>%
        pivot_wider(names_from = sostenimiento, values_from = n) %>%
        mutate(public = ifelse(is.na(public), 0, public)) %>%
        mutate(mixed = ifelse(is.na(mixed), 0, mixed)) %>%
        mutate(private = ifelse(is.na(private), 0, private)) %>%
        mutate(total = public + mixed + private) %>%
        mutate(public.percent = public/total) %>%
        mutate(mixed.percent = mixed/total) %>%
        mutate(private_percent = private/total)

df2 <- df1 %>%
        arrange(-private_percent) %>%
        mutate(private_percent = private_percent * 100) %>%
        select(canton, private_percent, total)

head(df2, 10)   
```

```{r}
df1 <- df %>%
        group_by(sostenimiento) %>%
        summarise(no.teachers = sum(total.docentes))
df1
```

```{r}
df1 <- df %>%
        group_by(sostenimiento) %>%
        count()
df1
```

```{r}
y2009s <- "https://educacion.gob.ec/wp-content/plugins/download-monitor/download.php?id=10555"
y2010s <- "https://educacion.gob.ec/wp-content/plugins/download-monitor/download.php?id=10557"
y2011s <- "https://educacion.gob.ec/wp-content/plugins/download-monitor/download.php?id=10559"
y2012s <- "https://educacion.gob.ec/wp-content/plugins/download-monitor/download.php?id=10561"
y2013s <- "https://educacion.gob.ec/wp-content/plugins/download-monitor/download.php?id=12176"
y2014s <- "https://educacion.gob.ec/wp-content/plugins/download-monitor/download.php?id=10567"
y2015s <- "https://educacion.gob.ec/wp-content/plugins/download-monitor/download.php?id=10569"
y2016s <- "https://educacion.gob.ec/wp-content/plugins/download-monitor/download.php?id=10573"
y2017s <- "https://educacion.gob.ec/wp-content/plugins/download-monitor/download.php?id=12645"
y2018s <- "https://educacion.gob.ec/wp-content/plugins/download-monitor/download.php?id=15711"
y2019s <- "https://educacion.gob.ec/wp-content/plugins/download-monitor/download.php?id=15713"

links1 <- list(y2009s, y2010s, y2011s, y2012s, y2013s, 
               y2014s, y2015s, y2016s, y2017s)
links2 <- list(y2018s, y2019s)
 
load.xlsx <- function(x, y, z){
        y <- read.xlsx(x, sheet = y, startRow = z)
        return(y)}

dat1 <- load.xlsx(y2009s, y= 1, z = 11)

dfB <- read.xlsx(y2018s, sheet = 2, startRow = 13)
dfC <- read.xlsx(y2019s,  sheet = "Registros Administrativos", startRow = 13)
listA <- lapply(links1, load.xlsx, y = 1, z = 11)

dfA <- do.call(rbind, listA)
names(dfA) <- replace_special_char(names(dfA))
dfA1 <- dfA %>%
        select(Periodo, Total.Estudiantes, Total.Docentes, 
               Nombre.Institucion, Zona.Inec, Sostenimiento, Canton, Provincia)

names(dfB) <- replace_special_char(names(dfB))
dfB1 <- dfB %>%
        select(Periodo, Total.Estudiantes, Total.Docentes, 
               Nombre.Institucion, Zona.Inec, Sostenimiento, Canton, Provincia)

names(dfC) <- replace_special_char(names(dfC))
names(dfC)[2] <- "duplicate"
dfC1 <- dfC %>%
        select(-duplicate) %>%
        select(Periodo, "Total.Estudiantes" = 'Total.estudiantes.(Inicial.a.3ro..De.Bachillerato)', 
               Total.Docentes, Nombre.Institucion, Zona.Inec, Sostenimiento, Canton, Provincia)

df <- rbind(dfA1, dfB1, dfC1)
saveRDS(df, "data.rds")
```

```{r}
df <- readRDS("data.rds")
df1 <- df %>%
        select("year" = Periodo, "tot.students" = Total.Estudiantes, 
               "tot.teachers" = Total.Docentes, Sostenimiento, "name" = Nombre.Institucion,
               "area" = Zona.Inec, Canton, Provincia) %>%
        mutate(year = str_extract(year, "^[0-9]{4}-[0-9]{4}"))  %>%
        #mutate(year = as.Date(paste0(year, "-01-01"))) %>%
        mutate(finance.source= factor(Sostenimiento, levels = c("Fiscal", "Fiscomisional", 
                                                                "Particular", "Municipal",
                                                                "Particular Laico",
                                                                "Particular Religioso"), 
                                      labels = c("public", "mixed", "private", "public",
                                                 rep("private", 2)))) %>%
        mutate(area = str_remove(area, "INEC")) %>%
        mutate(canton = replace_special_char(Canton)) %>%
        filter(tot.students != 0 & tot.teachers != 0)

df2 <- df1 #%>%
        #filter(!canton %in% c("SAMBORONDON", "GUAYAQUIL", "DURAN", "RUMINAHUI", "DAULE", "QUITO", "SALINAS"))

dfCS <- df2 %>%
        filter(finance.source == "public") %>%
        mutate(class.size = tot.students/tot.teachers) %>%
        group_by(year, canton, area) %>%
        summarise(class.size = median(class.size))

missing_to_zero <- function(var){return(ifelse(is.na(var), 0, var))}

df3<- df2 %>%
        group_by(year, finance.source, canton, area) %>%
        summarise(n = sum(tot.students)) %>%
        pivot_wider(names_from = finance.source, values_from = n) %>%
        mutate(public = missing_to_zero(public)) %>%
        mutate(mixed = missing_to_zero(mixed)) %>%
        mutate(private = missing_to_zero(private)) %>%
        mutate(prop = (private/sum(public, mixed, private))*100) %>%
        left_join(dfCS, by = c("year", "canton", "area")) %>%
        ungroup() %>%
        mutate(area = factor(area))

df4 <- df3 %>%
        filter(year %in% c("2010-2011", "2013-2014", "2016-2017", "2019-2020"))# %>%
        #filter(prop != 0)

p <- ggplot(dat = df4, aes(y = class.size, x = prop)) + theme_bw() 
p1 <- p + geom_point(alpha = 1/2) + facet_grid(area~year)
p2 <- p1 + geom_smooth(method = "lm", formula = y~x) 
p3 <- p2 + theme(axis.text.x = element_text(angle = 90)) + labs(
  x = "Proportion of students enrolled\nin private education (%)", 
  y = "Median of the average-class-size in public schools",
  title = "Relationship between private education attendance\nand public school class sizes in Ecuadorian cantons",
  subtitle = "Data reflects student enrolment and teacher positions at the start of the school year in all\nprimary and secondary education institutions. The average-class-size was first calculated\nfor each school. The median average-class-size was then taken at canton level.", 
  caption = "AMIE Data collected by INEC (2010-2019)")
p3

## This suggests that private institutions are concentrated in areas where class sizes are already large. 
df5 <- df4 %>%
  filter(area == "Urbana" & year == "2019-2020")
r <- lm(formula = prop ~ class.size, data = df5)
summary(r)
plot(r)

### The Bresusch-Pagan Test
## It involves using a variance function and using a χ2-test to test the null hypothesis that heteroskedasticity is not present (i.e. homoskedastic) against the alternative hypothesis that heteroskedasticity is present.

df5$resi <- r$residuals
var.func <- lm(resi^2 ~ class.size, data = df5)
summary(var.func)
qchisq(.95, df = 1)
```

```{r}
name_check <- function(x, y){
  for(i in 1:length(names(x))){
    if(isFALSE(identical(names(x)[i], names(y)[i]))){
      sapply(c(i, names(x)[i], names(y)[i]), print)}}
  message("all other colnames identical")}
```

```{r}
## Sources
"https://blogs.iadb.org/educacion/es/covid-19-y-la-transicion-de-la-educacion-privada-a-la-publica-en-ecuador/"
"https://www.elnuevosiglo.com.co/articulos/07-2020-la-transicion-de-la-educacion-privada-la-publica"
"https://www.infobae.com/america/mexico/2020/08/04/la-sep-esta-lista-para-migracion-de-alumnos-de-escuelas-privadas-a-publicas/"
"https://www.infobae.com/america/mexico/2020/08/04/la-sep-esta-lista-para-migracion-de-alumnos-de-escuelas-privadas-a-publicas/"
"https://www.24-horas.mx/2020/08/24/educacion-en-tiempos-de-covid-crece-migracion-de-escuelas-privadas-a-publicas/"
"https://www.eldiariodechihuahua.mx/local/temen-migracion-de-escuelas-privadas-a-publicas-20200805-1692951.html"
"https://www.milenio.com/politica/sep-lista-migracion-alumnos-escuelas-privadas-publicas"
"https://politica.expansion.mx/mexico/2020/08/10/el-covid-19-pone-en-riesgo-a-1-de-cada-4-escuelas-privadas"
"https://www.elfinanciero.com.mx/empresas/advierten-que-4-de-cada-10-de-escuelas-privadas-cerraran-por-educacion-a-distancia"
```


Context

Prop of priv/public schools, prop of priv/public teachers, boxplot class sizes. 
story: private schools have smaller class sizes than public schools. 

Inter American Bank data + Mexico
Story: migration from private schools to public schools. If situation does not change class sizes in public institutions will grow.

Is there historical evidence?
We expect it to be a static model instead of a lag model. 
Test for Autoregressive models of different orders. 

## Clean data set
```{r}
df <- readRDS("data.rds")
df1 <- df %>%
        select("year" = Periodo, "tot.students" = Total.Estudiantes, 
               "tot.teachers" = Total.Docentes, Sostenimiento, "name" = Nombre.Institucion,
               "area" = Zona.Inec, "canton" = Canton, "province" = Provincia) %>%
        mutate(year = str_extract(year, "^[0-9]{4}"))  %>%
        mutate(year = as.Date(paste0(year, "-05-01"))) %>%
        mutate(finance.source = factor(Sostenimiento, levels = c("Fiscal", "Fiscomisional", 
                                                                "Particular", "Municipal",
                                                                "Particular Laico",
                                                                "Particular Religioso"), 
                                      labels = c("public", "mixed", "private", "public",
                                                 rep("private", 2)))) %>%
        select(-Sostenimiento) %>%
        mutate(area = factor(str_remove(area, "INEC"), 
               levels = c("Rural", "Urbana"), labels = c("rural", "urban"))) %>%
        mutate(province = factor(province)) %>%
        #group_by(province) %>% ## Because some cantons in different provinces have the same names
        #mutate(canton = factor(replace_special_char(as.character(canton)))) %>%
        filter(tot.students != 0 & tot.teachers != 0) #%>%
        #ungroup()
```

```{r}
link <- "http://api.worldbank.org/v2/en/indicator/SE.XPD.TOTL.GD.ZS?downloadformat=csv"
temp <- tempfile()
download.file(link, temp, mode = "wb")
dfEDSP <- read.csv(unz(temp, filename = "API_SE.XPD.TOTL.GD.ZS_DS2_en_csv_v2_1345062.csv"), sep = ",", skip = 4)
unlink(temp)
```


# Country level time-series regression

## Step 1: Calculate Parameters at Country-level 
```{r}
df2 <- df1 %>%
        filter(finance.source == "public") %>%
        mutate(public.school.class.size = tot.students/tot.teachers) %>%
        group_by(year) %>%
        summarise(median.public.school.class.size = median(public.school.class.size))

missing_to_zero <- function(var){return(ifelse(is.na(var), 0, var))}

df3<- df1 %>%
        group_by(year, finance.source) %>%
        summarise(n = sum(tot.students)) %>%
        pivot_wider(names_from = finance.source, values_from = n) %>%
        mutate(public = missing_to_zero(public)) %>%
        mutate(mixed = missing_to_zero(mixed)) %>%
        mutate(private = missing_to_zero(private)) %>%
        mutate(proportion.private.students = (private/sum(public, mixed, private))*100) %>%
        left_join(df2, by = "year") %>%
        ungroup() 
```

step 2: Convert data to an object of class **ts**
```{r}
tsdf <- xts(x = df3[, 5:6], order.by = df3$year) # Number of observations per unit time
```

Step 3: Exploratory analysis of the data
The two  graphs suggest a negative correlation between the two variables. It must be acknowledged here that this correlation could be spurious. For example, Wooldridge stresses that any variables with a systematic negative and positive time trend can easily be misinterpreted to influence one another (p. 365). Such time trends can be accounted for by detrending or deseasonalising the data, i.e. by including time as an independent variable in the regression model. Nevertheless, time trends are plausibly irrelevant in this particular analysis; in the limited set of observations available, neither variable shows a clear linear or seasonal pattern over time. For this reason, I assume that the data need not be detrended nor deseasonalised. 

```{r}
par(mfrow = c(1,2))
plot(as.zoo(tsdf[, 1]),
     col = "blue",
     lwd = 2,
     ylab = "Proportion of Private Students",
     xlab = "Date",
     main = "Proportion of Ecuadorian learners\nenrolled in Private Schools",
     cex.main = 1)
plot(as.zoo(tsdf[, 2]),
     col = "red",
     lwd = 2,
     ylab = "Average-class-size at\nMedian Public School",
     xlab = "Date",
     main = "Class size in Public Schools",
     cex.main = 1)
```

```{r}
par(mfrow = c(1,2))
acf(tsdf[,1], lag.max = 12)
acf(tsdf[,2], lag.max = 12)
## For both variables autocorrelation until the first lag is statistically significant at a 95% confidence level. 
```
Step 4: create a model
```{r}
coe
```



```{r}

df

df3 <- df2 %>%
        filter(finance.source == "public") %>%
        mutate(class.size = tot.students/tot.teachers) %>%
        group_by(year, canton, area) %>%
        summarise(class.size = median(class.size))

missing_to_zero <- function(var){return(ifelse(is.na(var), 0, var))}

df4<- df2 %>%
        group_by(year, finance.source, canton, area) %>%
        summarise(n = sum(tot.students)) %>%
        pivot_wider(names_from = finance.source, values_from = n) %>%
        mutate(public = missing_to_zero(public)) %>%
        mutate(mixed = missing_to_zero(mixed)) %>%
        mutate(private = missing_to_zero(private)) %>%
        mutate(prop = (private/sum(public, mixed, private))*100) %>%
        left_join(dfCS, by = c("year", "canton", "area")) %>%
        ungroup() %>%
        mutate(area = factor(area))


p <- ggplot(dat = df4, aes(y = class.size, x = prop)) + theme_bw() 
p1 <- p + geom_point(alpha = 1/2) + facet_grid(area~year)
p2 <- p1 + geom_smooth(method = "lm", formula = y~x) 
p3 <- p2 + theme(axis.text.x = element_text(angle = 90)) + labs(
  x = "Proportion of students enrolled\nin private education (%)", 
  y = "Median of the average-class-size in public schools",
  title = "Relationship between private education attendance\nand public school class sizes in Ecuadorian cantons",
  subtitle = "Data reflects student enrolment and teacher positions at the start of the school year in all\nprimary and secondary education institutions. The average-class-size was first calculated\nfor each school. The median average-class-size was then taken at canton level.", 
  caption = "AMIE Data collected by INEC (2010-2019)")
p3

## This suggests that private institutions are concentrated in areas where class sizes are already large. 
df5 <- df4 %>%
  filter(area == "Urbana" & year == "2019-2020")
r <- lm(formula = prop ~ class.size, data = df5)
summary(r)
plot(r)

### The Bresusch-Pagan Test
## It involves using a variance function and using a χ2-test to test the null hypothesis that heteroskedasticity is not present (i.e. homoskedastic) against the alternative hypothesis that heteroskedasticity is present.

df5$resi <- r$residuals
var.func <- lm(resi^2 ~ class.size, data = df5)
summary(var.func)
qchisq(.95, df = 1)
```

```{r}
GDPGR_AR2 <- dynlm(ts(GDPGR_level) ~ L(ts(GDPGR_level)) + L(ts(GDPGR_level), 2))
```

Why is it bad?
Private schools concentrated in particular areas (inter American bank data). Furthermore, private schools tend to concentrate in areas where there public school classes are large. 




